Dynamic Species Interactions in the San Nicolas Island Kelp Forest

### 2018-07-04 EDM_SNI_analysis.Rmd
experimental analyses at the end

  # Model output-- time, observations, predictions. time is truncated, so re-expand it to fit with original periods
  t <- best.mod$model_output[[1]]$time

  # Save interaction coefficients
  coeffs <- best.mod$smap_coefficients[[1]]
  names(coeffs) <- c(causalvars,"const")
  coeffs <- coeffs %>% mutate(time=t)%>%complete(time=full_seq(x=c(1,630),1)) %>% mutate(period=rep(1:63,10))

  # long form for plotting
  coeffs.long <- coeffs %>%
    gather(key=effect,value=value,-period,-time,na.rm=T) %>%
    left_join(namekey, by=c("effect"="short"))

@@

  outlist <- list(model=best.mod,coeffs=coeffs,coeffs.long= coeffs.long,plot=int.box)
  return(outlist)
}
```

@@

```{r phys coeffs,echo=F,message=F,warning=F,fig.height=8}
# with boxplot
all.phys.coeffs.plot <- all.mod.coeffs.long %>%
  filter(effect %in% phys.vars)%>%
  mutate(label = case_when(
    mod == "mac" ~ "Macrocystis >1m",
    mod == "ymac" ~ "Macrocystis <1m",
    mod == "lam" ~ "Laminaria",
    mod == "pter" ~ "Pterygophora",
    mod == "red" ~ "Red urchin",
    mod == "purp" ~ "Purple urchin")
  ) %>%
  ggplot(aes(effect,value))+
  scale_x_discrete(labels=c(
    "mei"="MEI",
    "npgo"="NPGO",
    "pdo"="PDO",
    "waves"="SWH",
    "sst"="SST")
    )+
  geom_boxplot(fill="dodgerblue2",outlier.alpha=0.4,position = "identity")+
  stat_summary(fun.y=mean, colour="white", geom="point", 
               shape=17, size=2,show.legend = FALSE)+
  geom_hline(yintercept=0,linetype=2)+
  ylim(-2,2)+
  coord_flip()+
  labs(y="Estimated Interaction Coefficient",x="Physical Forcing")+
  facet_grid(mod_plotting_name~.)+
  geom_text(aes(x=1.5,y=2,label=label),hjust=1,family="Rockwell",fontface="italic",check_overlap = TRUE,size=4,show.legend = FALSE)+
  theme(strip.background = element_blank(),
        strip.text = element_blank())
all.phys.coeffs.plot
# ggsave("plots/phys_coeffs.png",plot=all.phys.coeffs.plot)
```

@@

```{r individual_effects,fig.width=7}

phys_effects_on_interactions <- purrr::map(as.list(study_spp),function(spp) {
  quants <- c(0.05,0.5,0.95)
  plot_title <- namekey$long[match(spp,namekey$short)]
  all.coeffs.with.phys %>%
  filter(!is.na(type),mod==spp)%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
    geom_point(col='gray60',alpha=0.6)+
    geom_hline(yintercept = 0,linetype=2)+
    geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
    scale_color_manual(values=c('mac'="navyblue",'ymac'="gray30",'pter'="darkgreen",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),name="effect")+
    labs(title=paste("Physical forcing effect on interactions with",plot_title),x="Normalized Index Value",y="Interaction Strength")+
    facet_grid(type~external_forcing,scales = "free_y")+
    theme(strip.text=element_text(size=8),
          legend.text = element_text(size=12),
          legend.position="right",
          legend.margin=margin(t = -0.5, unit='line'),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"),
        )
})
ggsave("plots/phys_effects_red.png",plot=phys_effects_on_interactions[[1]])
ggsave("plots/phys_effects_purp.png",plot=phys_effects_on_interactions[[2]])
ggsave("plots/phys_effects_lam.png",plot=phys_effects_on_interactions[[3]])
ggsave("plots/phys_effects_pter.png",plot=phys_effects_on_interactions[[4]])
ggsave("plots/phys_effects_mac.png",plot=phys_effects_on_interactions[[5]])
ggsave("plots/phys_effects_ymac.png",plot=phys_effects_on_interactions[[6]])

@@

```{r some single plots}
# macrocystis, pterygophora, and temperature
quants<-c(0.05,0.5,0.95)
all.coeffs.with.phys %>%
  filter(type=="algal competition",mod%in% c('mac','pter'),effect %in% c('mac','pter'),external_forcing=="sst")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
   geom_point(col='gray60',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="white")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized SST",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
  plot_theme_black+
    theme(strip.text=element_blank(),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

# macrocystis, pterygophora, and NPGO
all.coeffs.with.phys %>%
  filter(type=="algal competition",mod%in% c('mac','pter'),effect %in% c('mac','pter'),external_forcing=="npgo")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
   geom_point(col='gray60',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="white")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized NPGO",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
  plot_theme_black+
    theme(strip.text=element_blank(),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

# macrocystis, pterygophora, herbivory, and NPGO
all.coeffs.with.phys %>%
  filter(type=="herbivory",mod%in% c('mac','pter'),effect %in% c('red','purp'),external_forcing=="npgo")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
   geom_point(col='gray60',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="white")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized NPGO",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
  plot_theme_black+
    theme(strip.text=element_blank(),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

# macrocystis, pterygophora, and waves
all.coeffs.with.phys %>%
  filter(type=="algal competition",mod%in% c('mac','pter'),effect %in% c('mac','pter'),external_forcing=="waves")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
   geom_point(col='gray60',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="white")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized SWH",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
  plot_theme_black+
    theme(strip.text=element_blank(),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

@@

### Effect of Macrocystis on Laminaria/Pterygophora interaction
```{r disturbance and mac,fig.height=7,fig.width=5}
quants <- c(0.05,0.5,0.95)
mac.lam.pter <- westend %>%
  select(one_of(c("mac","period"))) %>%
  distinct() %>%
  right_join(all.mod.coeffs.long,by="period")%>%
  gather("external_spp","ext_value",mac)%>%
  filter(effect=='pter',mod=='lam')%>%
  ggplot(aes(ext_value,value))+
  geom_point(col='gray60',alpha=0.6)+
  geom_hline(yintercept = 0,linetype=2)+
  geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
  labs(x="Normalized Density",y="Effect of Pterygophora on Laminaria")
mac.lam.pter
```
## References

>>>>

  # gather and organize model output
  t <- best.mod$model_output[[1]]$time

  coeff.names <- c(causalvars,"const")
  # Save interaction coefficients
  coeffs <- best.mod$smap_coefficients[[1]]
  names(coeffs) <- coeff.names

  # # Model output-- time, observations, predictions. match time indicator with observations (one period ahead)
  coeffs <- coeffs %>% mutate(time=t)%>%complete(time=full_seq(x=c(1,630),1)) %>% mutate(period=rep(1:63,10))

  # Save coefficient variances (sensu Deyle 2016; we have a nxn variance-covariance matrix for each estimated point, where n is the total number of causal variables for whom coefficients are estimated, plus the constant term)
  # Variances for individual coefficients are the diagonals of each covariance matrix
  coeffs.variances <- best.mod$smap_coefficient_covariances[[1]] %>%
    purrr::map_dfr(function(x) {
      # if no estimate, no variance associated with it (NULL)
      if(is.null(x)) x <- matrix(NaN,nrow=length(coeff.names),ncol=length(coeff.names))
      # pull out diagonal elements, add variable names
      out <- t(diag(x)) %>% as.data.frame()
      names(out) <- coeff.names
      out
      })%>%

    # add period identifier and overall time stamp
    # mutate(time=t) %>% complete(time=full_seq(c(1,630),1)) %>% mutate(period=rep(1:63,10))
    mutate(time=t)%>%complete(time=full_seq(x=c(1,630),1)) %>% mutate(period=rep(1:63,10))

  # long form for plotting, including variance
  coeffs.long <- coeffs %>%
    gather(key=effect,value=value,-period,-time,na.rm=T) %>%
    left_join(namekey, by=c("effect"="short"))

  # long form variances
  coeffs.variances.long <- coeffs.variances %>%
    gather(key=effect,value=variance,-period,-time,na.rm=T) %>%
    left_join(namekey,by=c("effect"="short"))

  coeffs.long <- coeffs.long %>%
    left_join(coeffs.variances.long,by=c("time","period","effect","dataset","long","plotting")) %>%
    # add standard deviation and +/- 95% CI (1.96*SD)
    mutate(sd=sqrt(variance),upper=value+1.96*sd,lower=value-1.96*sd)

@@

  outlist <- list(model=best.mod,coeffs=coeffs,coeffs.variances=coeffs.variances,coeffs.long= coeffs.long,plot=int.box)
  return(outlist)
}
```

@@

```{r phys coeffs,echo=F,message=F,warning=F,fig.height=8,fig.width=8}
# with boxplot
all.phys.coeffs.plot <- all.mod.coeffs.long %>%
  filter(effect %in% phys.vars)%>%
  mutate(label = case_when(
    mod == "mac" ~ "Macrocystis >1m",
    mod == "ymac" ~ "Macrocystis <1m",
    mod == "lam" ~ "Laminaria",
    mod == "pter" ~ "Pterygophora",
    mod == "red" ~ "Red urchin",
    mod == "purp" ~ "Purple urchin")
  ) %>%
  ggplot(aes(effect,value))+
  scale_x_discrete(labels=c(
    "mei"="MEI",
    "npgo"="NPGO",
    "pdo"="PDO",
    "waves"="SWH",
    "sst"="SST")
    )+
  geom_boxplot(fill="dodgerblue2",outlier.alpha=0.4,position = "identity")+
  stat_summary(fun.y=mean, colour="white", geom="point", 
               shape=17, size=2,show.legend = FALSE)+
  geom_hline(yintercept=0,linetype=2)+
  ylim(-2,2)+
  coord_flip()+
  labs(y="Estimated Interaction Coefficient",x="Physical Forcing")+
  facet_grid(mod_plotting_name~.)+
  geom_text(aes(x=1.5,y=2,label=label),hjust=1,family="Rockwell",fontface="italic",check_overlap = TRUE,size=4,show.legend = FALSE)+
  theme(strip.background = element_blank(),
        strip.text = element_blank())
all.phys.coeffs.plot
# ggsave("plots/phys_coeffs.png",plot=all.phys.coeffs.plot)
```

@@

```{r individual_effects,fig.width=7}
phys_effects_on_interactions <- purrr::map(as.list(study_spp),function(spp) {
  quants <- c(0.05,0.5,0.95)
  plot_title <- namekey$long[match(spp,namekey$short)]
  all.coeffs.with.phys %>%
  filter(!is.na(type),mod==spp)%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
    geom_linerange(aes(ymin=lower,ymax=upper),col='gray60')+
    geom_point(col='gray60',alpha=0.6)+
    geom_hline(yintercept = 0,linetype=2)+
    geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
    scale_color_manual(values=c('mac'="navyblue",'ymac'="gray30",'pter'="darkgreen",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),name="effect")+
    labs(title=paste("Physical forcing effect on interactions with",plot_title),x="Normalized Index Value",y="Interaction Strength")+
    facet_grid(type~external_forcing,scales = "free_y")+
    theme(strip.text=element_text(size=8),
          legend.text = element_text(size=12),
          legend.position="right",
          legend.margin=margin(t = -0.5, unit='line'),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line")
        )
})
ggsave("plots/phys_effects_red.png",plot=phys_effects_on_interactions[[1]])
ggsave("plots/phys_effects_purp.png",plot=phys_effects_on_interactions[[2]])
ggsave("plots/phys_effects_lam.png",plot=phys_effects_on_interactions[[3]])
ggsave("plots/phys_effects_pter.png",plot=phys_effects_on_interactions[[4]])
ggsave("plots/phys_effects_mac.png",plot=phys_effects_on_interactions[[5]])
ggsave("plots/phys_effects_ymac.png",plot=phys_effects_on_interactions[[6]])

@@

```{r some single plots}
# macrocystis, pterygophora, and temperature
quants<-c(0.05,0.5,0.95)
all.coeffs.with.phys %>%
  filter(type=="algal competition",mod%in% c('mac','pter'),effect %in% c('mac','pter'),external_forcing=="sst")%>%

  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
    geom_linerange(aes(ymin=lower,ymax=upper),col='gray60')+
    geom_point(col='gray60',alpha=0.6)+
    geom_hline(yintercept = 0,linetype=2,col="white")+
    geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
    scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
    labs(x="Normalized SST",y="Interaction Strength")+
    facet_grid(~mod,scales = "free_y")+
  plot_theme_black+
    theme(strip.text=element_blank(),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

# macrocystis, pterygophora, and NPGO
all.coeffs.with.phys %>%
  filter(type=="algal competition",mod%in% c('mac','pter'),effect %in% c('mac','pter'),external_forcing=="npgo")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
  geom_linerange(aes(ymin=lower,ymax=upper),col='gray60')+ 
  geom_point(col='gray60',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="white")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized NPGO",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
  plot_theme_black+
    theme(strip.text=element_blank(),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

# macrocystis, pterygophora, herbivory, and NPGO
all.coeffs.with.phys %>%
  filter(type=="herbivory",mod%in% c('mac','pter'),effect %in% c('red','purp'),external_forcing=="npgo")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
   geom_linerange(aes(ymin=lower,ymax=upper),col='gray60')+
    geom_point(col='gray60',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="white")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized NPGO",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
  plot_theme_black+
    theme(strip.text=element_blank(),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

# macrocystis, pterygophora, and waves
all.coeffs.with.phys %>%
  filter(type=="algal competition",mod%in% c('mac','pter'),effect %in% c('mac','pter'),external_forcing=="waves")%>%
  ggplot(aes(ext_value,value,col=factor(effect,levels=c("mac","ymac","pter","lam","purp","red"))))+
  geom_linerange(aes(ymin=lower,ymax=upper),col='gray60')+
  geom_point(col='gray60',alpha=0.6)+
   geom_hline(yintercept = 0,linetype=2,col="white")+
   geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
   scale_color_manual(values=c('mac'="lightblue",'ymac'="gray30",'pter'="red",'lam'="darkcyan",'purp'="mediumpurple4",'red'="darkred"),guide="none")+
   labs(x="Normalized SWH",y="Interaction Strength")+
   facet_grid(~mod,scales = "free_y")+
  plot_theme_black+
    theme(strip.text=element_blank(),
          plot.margin=unit(c(0.5, 0.5, 0.5, 0.5), units="line"))

@@

### Effect of Macrocystis on Laminaria/Pterygophora interaction
```{r disturbance and mac,fig.height=7,fig.width=5}
quants <- c(0.05,0.5,0.95)
mac.lam.pter <- westend %>%
  select(one_of(c("mac","period"))) %>%
  distinct() %>%
  right_join(all.mod.coeffs.long,by="period")%>%
  gather("external_spp","ext_value",mac)%>%
  filter(effect=='pter',mod=='lam')%>%
  ggplot(aes(ext_value,value))+
  geom_point(col='gray60',alpha=0.6)+
  geom_hline(yintercept = 0,linetype=2)+
  geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
  labs(x="Normalized Density",y="Effect of Pterygophora on Laminaria")
mac.lam.pter
```
### Effect of urchin density on Macrocystis/Pterygophora interaction

```{r urchin dens mac pter}
quants <- c(0.05,0.5,0.95)
urchin.mac.pter <- westend %>%
  select(one_of(c("red","period"))) %>%
  distinct() %>%
  mutate(time=row_number())%>%
  right_join(all.mod.coeffs.long,by=c("period",'time'))%>%
  gather("external_spp","ext_value",red)%>%
  filter(effect %in% c('pter','mac'),mod %in% c('pter','mac'),!is.na(ext_value))%>%
  ggplot(aes(ext_value,value))+
  geom_point(col='gray60',alpha=0.6)+
  geom_hline(yintercept = 0,linetype=2)+
  geom_quantile(quantiles=quants,lineend = "round",size=1,formula=y~x)+
  labs(x="Normalized Red Urchin Density",y="Interaction Strength")+
  facet_wrap(~mod)
urchin.mac.pter
```

### Do conditions affect average interaction strength?

```{r conditions and interaction strength}
# disturbance. top quartile vs. bottom quartile
all.coeffs.with.phys %>%
  mutate(av_value=abs(value))%>%
  filter(external_forcing=='pdo', type=='algal competition', ext_value>=quantile(westend$pdo,0.75,na.rm=T)|ext_value<=quantile(westend$pdo,0.25,na.rm=T))%>%
  mutate(stress_level=ifelse(ext_value>=quantile(westend$pdo,0.75,na.rm=T),'high','low'))%>%
  ggplot(aes(av_value,..density..,fill=stress_level))+
  geom_density(alpha=0.6)+
  geom_vline(xintercept = 0,linetype=2,color='black')+
  labs(x='Interaction Strength',y='Probability Density')

## References
